---
title: "Marketing mix modelling case study"
author: 'Gabriel Scozzarro'
date: '17/05/2021'
output: hrbrthemes::ipsum_pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  dev = "cairo_pdf")
```

## 1.0 Introduction

Understanding the ROI across all of your paid marketing channels is a top priority for senior-level executives across every industry and every geographical market. Getting a clear 
sense of the ROI on each channel allows companies to answer really important questions about performance changes increasing a specific marketing channel budget, marketing channel 
saturation or budget optimization across different marketing channel.

### 1.1 Scope of work & Business task

This case study, aim to showcase how to use marketing campaign data to create a marketing mix modeling solution. This work provide also a budget prediction to reach a 
predetermined business goal.

```{r Tools, warning = FALSE, error = FALSE, message = FALSE, echo = FALSE}
library(tidyverse)
library(scales)
library(gridExtra)
library(skimr)

library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
library(kableExtra)


extrafont::loadfonts()
```


## 2.0 Data
The data set used includes information about several different marketing campaign performed using the email channel.
```{r Data import, warning = FALSE, error = FALSE, message = FALSE, echo = FALSE}
email_mk_data<- read.csv("Data/marketing_data.csv",header = TRUE, sep = ";")

#fixing ----
colnames(email_mk_data)[1] <- "Campaign"

head(email_mk_data, 5) %>% kable('latex', digits = 10, caption = 'Marketing campaign data sample', booktabs = T) %>% 
                           kable_styling(full_width = TRUE, font_size = 9, latex_options = c("striped",'scale_down'))
```

\pagebreak

The data set was aggregated to the campaign level, meaning each row represents one unique campaign.
The Return value need to some measure of value to your business and in this case was revenue generated.
Return value can vary depending on:

* __What is valuable for your business__: Are you in the retail industry where customers are making purchases directly on your site? If so, you may want to use Revenue or Orders 
for your measure of value. Are you a media site where consumption of content is the most important metric? If so, you may want to use blog views or subscriptions as your measure 
of value. Return must always be specific to your industry and your business organization.

* __What type of attribution model the business use__: The amount of Return allocated to each campaign will be directly affected by the attribution method that you deploy. 
Most companies are using simple, rules-based attribution methods like first touch or last touch.

The Spend value is clearly straight forward, it provide information about the budget spent on the campaign.


## 3.0 Result
Assuming a business goal of $600.000 in revenue from those email marketing campaign, our model suggested a total budget incremented of around 53% for an amount of
$171.455


## 4.0 Analysis

Since data was already aggregated at high level, we can jump directly to visualize the campaigns performance.

```{r Analysis 1 , warning = FALSE, error = FALSE, message = FALSE, echo = FALSE, out.width= '50%'}

email_mk_data %>% ggplot(aes(Campaign, Spend, fill = Campaign)) +
                  geom_bar(stat = "identity") +
                  scale_y_continuous(labels = dollar) +
                  scale_fill_brewer(palette = "Set3") +
                  theme_ipsum() +
                  theme(axis.text.x = element_text(angle = 45, hjust = 0.75), 
                        legend.position = "none") +
                  xlab("") +
                  labs(title = "Invesment in different marketing campaign")

email_mk_data %>% ggplot(aes(Campaign, Return, fill = Campaign)) +
                  geom_bar(stat = "identity") +
                  scale_y_continuous(labels = dollar) +
                  scale_fill_brewer(palette = "Set3") +
                  theme_ipsum() +
                  theme(axis.text.x = element_text(angle = 45, hjust = 0.75), 
                        legend.position = "none") +
                  xlab("") +
                  labs(title = "ROI in different marketing campaign")

```
Buy one get one was the most expensive campaign and it had a return of 3x on the investment. The semi-annual sale was the most effective campaign with a return of  around 4x on 
the investment. 
If we confront the Spend and Return value across all the email campaign, the relationship between them, at a first glance, might be interpreted as linear model.

```{r Analysis 2 , warning = FALSE, error = FALSE, message = FALSE, echo = FALSE, out.width= '50%'}

email_mk_data %>% ggplot(aes(Spend, Return)) +
                  geom_point(size = 2) +
                  geom_smooth(method = "lm") +
                  scale_x_continuous(labels = dollar) +
                  scale_y_continuous(labels = dollar) +
                  theme_ipsum() +
                  labs(title = "Email campaign Spent - Return relationship")
```

Using a linear model in this case raise up 2 problems:

* A linear model assumes that you have infinite growth.

* There are real-world phenomena like market saturation and email fatigue that suggest infinite growth is not actually possible.

To correctly model the email campaign data, a model that exhibits diminishing marginal returns is needed. The __ADBUDG model__ is a very flexible model that incorporates diminishing 
marginal returns. The ADBUDG model is defined as follows:

$$Return = B + (A-B)   \frac{\mathrm{Spend^C} }{\mathrm{D +Spend^C}}$$

Where:

* A – The maximum amount of return possible for a campaign given a long term investment.
* B – The minimum amount of return possible for a campaign given a long term investment.
* C – Controls the shape of the curve.
* D – Represents initial market share or market saturation effects.

\pagebreak
Using this model the result is:

```{r Analysis 3 , warning = FALSE, error = FALSE, message = FALSE, echo = FALSE, out.width= '80%', fig.align='center'}

#ADBUDG model

Ufun<-function(x, Spend, Return) {
  predictedReturn = x[2] + (x[1] - x[2])*((Spend^x[3])/(x[4] + (Spend^x[3])))
  errorSq = (predictedReturn - Return)^2
  sumSqError = sum(errorSq)
  return(sumSqError)
}
# The “x” variable is a vector of length four and represents the four parameters 
# of the ADBUDG function.  Spend and Return are also vectors which correspond with 
# the Spend and Return data from your data file.  Their length will be the same as 
# he number of data points in your data file.  The temp1 object calculates the predicted 
# Return from the ADBUDG equation.  That vector is then used in the next line when we 
# calculate the squared error for our predicted values, which are stored in the temp2 object.
# Lastly, we sum the temp2 object to get a single value for the sum of squared errors, which 
# is stored in the temp3 object.  The temp3 object is the return object for this function.  
# When we minimize this function, we are essentially minimizing the sum of squared error, 
# which is actually the same thing that is done in linear regression.
startValVec = c(25000,100,1.5,100000)
minValVec = c(0,0,1.01,0)
maxValVec = c(500000, 500000, 2, 10000000)

optim.parms<-nlminb(objective=Ufun,start=startValVec,
                    lower=minValVec,
                    upper=maxValVec,
                    control=list(iter.max=1000000,eval.max=20000),
                    Spend = email_mk_data$Spend,
                    Return = email_mk_data$Return)

a = optim.parms$par[1]
b = optim.parms$par[2]
c = optim.parms$par[3]
d = optim.parms$par[4]

curveDFx = seq(from=0, to=max(email_mk_data$Spend)*2, length.out=10000)
curveDFy = b+(a-b)*((curveDFx^c)/(d+(curveDFx^c)))
curveDF = data.frame(Spend = curveDFx, Return = curveDFy)

maxX = 1.05*max(curveDFx, max(email_mk_data$Spend))
maxY = 1.05*max(curveDFy, max(email_mk_data$Return))

myPlotDataDF = data.frame(Return = email_mk_data$Return, Spend = email_mk_data$Spend)
optimLineDF = data.frame(Spend = curveDFx, Return = curveDFy)

myPlotDataDF %>% ggplot(aes(Spend, Return)) +
                 geom_point(size = 2) +
                 geom_line(data = optimLineDF, aes(x = Spend, y = Return, color = "darkgreen")) +
                 scale_color_manual(labels = "Optimized ADBUDG Fit",values=c('darkgreen')) +
                 coord_cartesian(ylim = c(0,maxY), xlim = c(0,maxX)) +
                 scale_x_continuous(labels = dollar) +
                 scale_y_continuous(labels = dollar) +
                 theme_ipsum() +
                 theme(legend.title=element_blank(), legend.position = "bottom") +
                 labs(title = "Email campaign ADBUDG model")

```

As we can see the green line which represents the ADBUDG model perfectly fit our value and exhibits the diminishing marginal returns effect.
Now that we have a good model we can estimate how much the spent budget should be increased in each campaigns to reach that business goal.

Assuming the business want to reach a goal of $600.000 in return, the estimated budget for each campaign are as fallow:

```{r Analysis 4 , warning = FALSE, error = FALSE, message = FALSE, echo = FALSE, out.width= '85%', fig.align='center'}

adbudgReturn = function(a,b,c,d,Spend){
  adbudgReturn = sum(b+(a-b)*((Spend^c)/(d+(Spend^c))))
  return(adbudgReturn)
}

returnGoal = 600000
increment = 1000
oldSpendVec = email_mk_data$Spend
oldReturn = adbudgReturn(a,b,c,d,oldSpendVec)
newSpendVec = oldSpendVec

totalSpend = sum(oldSpendVec)
totalReturn = oldReturn                 

while(totalReturn < returnGoal){
  incReturns = NULL
  for(i in 1:length(oldSpendVec)){
    oldSpendTemp = newSpendVec[i]
    newSpendTemp = newSpendVec[i] + increment
    
    oldReturnTemp = b+(a-b)*((oldSpendTemp^c)/(d+(oldSpendTemp^c)))
    newReturnTemp = b+(a-b)*((newSpendTemp^c)/(d+(newSpendTemp^c)))
    
    incReturns[i] = newReturnTemp - oldReturnTemp
    
  }
  
  winner = which.max(incReturns)
  newSpendVec[winner] = newSpendVec[winner] + increment
  
  totalSpend = totalSpend + increment
  totalReturn = adbudgReturn(a,b,c,d,newSpendVec)
  
}

newReturnVec = b+(a-b)*((newSpendVec^c)/(d+(newSpendVec^c)))
myRecommendedData = data.frame(Campaign = email_mk_data$Campaign,
                               Channel = email_mk_data$Channel,
                               Return = newReturnVec,
                               Spend = newSpendVec)

reco_bud<- sum(myRecommendedData$Spend) # Recommended Spend
roi_est<- sum(myRecommendedData$Return)  # Estimated Return from Recommended Spend
perc_budget_recoinc<- sum(myRecommendedData$Spend)/sum(email_mk_data$Spend) - 1  # % Increase in Spend to get $600K


compareDF = data.frame(Campaign = rep(email_mk_data$Campaign,2), 
                       spendType = rep(c("Actual Spend","Recommended Spend"), 
                                       each=dim(email_mk_data)[1]), 
                       Spend = c(email_mk_data$Spend, myRecommendedData$Spend))


budget_var_plot<- compareDF %>% ggplot(aes(Campaign, Spend, fill = spendType)) +
                                geom_bar(stat = "identity", position=position_dodge()) +
                                scale_fill_brewer(palette = "Set1", name = "") +
                                scale_y_continuous(labels = dollar) +
                                theme_ipsum() +
                                theme(axis.text.x = element_text(angle = 45, hjust = .80)) +
                                labs(title = "Campaign investments", 
                                     subtitle = "Actual vs recommended buget to reach goal")

budget_var_plot
```